Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2023-01-11 supplement.tex
clarify optimization

For each model tested, we ran multiple rounds of likelihood optimization,
alternating between the \emph{optimize\_log\_fmin},
\emph{optimize\_log\_powell}, and \emph{optimize\_lbfgsb} methods to explore
parameter space and hone in on the best fit parameters.  We imposed bounds on
many of the parameters in the fits (most of which did not hit the imposed
bounds). Population sizes had a lower bound of $N_e=100$, migration rates were
bounded between 0 and $10^{-3}$, and admixture proportions between 0 and 1.
Event times were only constrained to occur in the proper order (so that
populations were required to have positive existence times, for example).
Initial guesses for parameters were chosen from demographically plausible
starting points and then perturbed to explore space, using gradient descent (on
the log of the parameters). The best fits from these initial rounds of
optimization were then chosen as the starting points for optimization using the
Powell and/or the L-BFGS-B methods (as implemented in the SciPy optimization
package \citep{Virtanen2020-kr}). This process was repeated with alternating
optimization methods until the best fit parameter set converged consistently.

>>>>

For each model tested, we ran multiple rounds of likelihood optimization,
alternating between the \emph{optimize\_log\_fmin},
\emph{optimize\_log\_powell}, and \emph{optimize\_lbfgsb} methods to explore
parameter space and hone in on the best fit parameters.  
We imposed bounds on many of the parameters in the fits.
This is done primarily to avoid optimisation exploring parts of the parameter space 
that are unrealistic (e.g., negative migration rates), where the model calculation can 
be costly (e.g., population splits billions of years ago), and where model assumptions 
can break down (e.g., very small populations).  Population sizes had a lower bound of 
$N_e=100$, migration rates were
bounded between 0 and $10^{-3}$, and admixture proportions between 0 and 1.
Event times were only constrained to occur in the proper order (so that
populations were required to have positive existence times, for example).
Most best-fit parameters were away from these boundaries.

Initial guesses for parameters were chosen from demographically plausible
starting points and then perturbed to explore space, using gradient descent (on
the log of the parameters). The best fits from these initial rounds of
optimization were then chosen as the starting points for optimization using the
Powell and/or the L-BFGS-B methods (as implemented in the SciPy optimization
package \citep{Virtanen2020-kr}). This process was repeated with alternating
optimization methods until the best fit parameter set converged consistently.
